Online Tuning of EASY-Backfilling using Queue Reordering Policies.

### 2018-01-08 paper/src/paper.tex
changes to the beginning of the paper

%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Online (Bandit) Policy Selection for EASY-Backfilling}

@@

  The EASY-FCFS heuristic is the basic building block of job scheduling
  policies in most parallel High Performance Computing platforms. Despite its
  simplicity, and the guarantee of no job starvation, it could still be
  improved on a per-system basis. Such tuning is difficult because of
  non-linearities in the scheduling process. The study conducted in this paper
  considers an online approach to the automatic tuning of the EASY heuristic
  for HPC platforms.  More precisely, we consider the problem of selecting a
  reordering policy for the job queue under several feedback modes. We show via
  a comprehensive experimental validation on actual logs that noisy feedback
  (using a simulator, which is necessarily imprecise) recovers existing
  \textit{in-hindsight} results that allow to divide the average waiting time
  by almost 2. Moreover, we show that bandit feedback can be used by an
  Epsilon-greedy multi-armed bandit algorithm to cut the average waiting time
  by 40\% without using a simulator.

@@

In order to simulate how the simulation strategy for selecting policies would
work on non simulated traces, we randomly introduce noise in the estimate of
the policy cost defined by Eq.~\eqref{eq:cost-exact} by rescaling it, either
down or up:
%
\begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) =
 \rho \sum_{j=1}^{l(\Delta_t)} \, \mbox{Wait}_j^{P_i} \end{equation}
%
where $\rho$ is uniformly sampled in the interval $[0.80,1.20]$, thus adding a
+/- 20\% noise on the waiting time estimated by the simulator. The overall cost
is then defined in the same way as before, leading to:
%
\begin{equation} \label{eq:tot-cost-noisy} w(\rightarrow T;P_i) =
\sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i) \end{equation}
%
As before (see Eq.~\ref{eq:select-exact}), the policy that minimizes the above
cost is selected for the period $\Delta_T$.

@@

The experimental validation performed in this work uses a mixture of small and
large systems from different periods. Table~\ref{tab:logs} presents the 7 logs,
which can all be obtained from the Parallel Workload
Archive~\cite{Feitelson20142967}. We use the ``cleaned''\cite{pwa-clean} version
of the logs as per the archive. We impose an additional filtering step to the
workloads\footnote{This filtering step is available in the reproducible
workflow\cite{repro} as shell script
\lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
clean erroneous data:

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{jsm}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. In our experience, randomly
shuffling months in a trace at random is enough to generate contradictory
conclusions about which algorithm is better. This is a known
problem~\cite{flurries} that motivated decades of research in workload
modeling~\cite{feitbook}. A recent approach~\cite{feitresampling} suggests to
resample data directly from the original traces. In this work, we use a simple
implementation of these ideas.  More precisely, we want to sample the job
submission process, which is achieved in the following manner: (a) For every
week in a resampled trace, we iterate over every existing user in the original
trace; (b) For each of these users, we choose a week of the original trace
uniformly at random; (c) The jobs from this user in this original week are
chosen to be present in the week being resampled.  This simple mechanism
enables us to preserve the dependency within a week.  Although the system state
may not be independent from one week to an other, we rely on existing
work~\cite{jsspp17} and assume for these experiments that the system is in a
stationary state. While we do provide guidelines as how to incorporate tracking
variant of our algorithms in Section~\ref{sec:online} by introducing a decaying
discount factor $\lambda$, this will not be studied in the experiments where 
$\lambda=1$, and is left for future research.

>>>>

%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Online Tuning of EASY-Backfilling using Queue Reordering Policies.}

@@

  The EASY-FCFS heuristic is the basic building block of job scheduling
  policies in most parallel High Performance Computing platforms. Despite its
  simplicity, and the guarantee of no job starvation, it could still be
  improved on a per-system basis. Such tuning is difficult because of
  non-linearities in the scheduling process. The study conducted in this paper
  considers an online approach to the automatic tuning of the EASY heuristic
  for HPC platforms.  More precisely, we consider the problem of selecting a
  reordering policy for the job queue under several feedback modes. We show via
  a comprehensive experimental validation on actual logs that periodic
  simulation of historical data can be used to recover existing
  \textit{in-hindsight} results that allow to divide the average waiting time
  by almost 2. This results holds even when the simulator results are noisy.
  Moreover, we show that good performances can still be obtained without a
  simulator, under what is called bandit feedback - when we can only observe
  the performance of the algorithm that was picked on the live system.  Indeed,
  a simple multi-armed bandit algorithm can reduce the average waiting
  time by 40\%.

@@

In order to simulate how the simulation strategy for selecting policies would
work on non simulated traces, we randomly introduce noise in the estimate of
the policy cost defined by Eq.~\eqref{eq:cost-exact} by rescaling it, either
down or up:
%
\begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) =
 \rho_t^i \sum_{j=1}^{l(\Delta_t)} \, \mbox{Wait}_j^{P_i} \end{equation}
%
where $\rho_t^i$ is uniformly sampled in the interval $[0.80,1.20]$, thus adding a
+/- 20\% multiplicative noise on the waiting time estimated by the simulator. The overall cost
is then defined in the same way as before, leading to:
%
\begin{equation} \label{eq:tot-cost-noisy} w(\rightarrow T;P_i) =
\sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i) \end{equation}
%
As before (see Eq.~\ref{eq:select-exact}), the policy that minimizes the above
cost is selected for the period $\Delta_T$.

@@

The experimental validation performed in this work uses a mixture of small and
large systems from different periods. These platforms are all targeted by this
work, since they all are homogenous systems that run a batch scheduler.
Table~\ref{tab:logs} presents the 7 logs, which can all be obtained from the
Parallel Workload Archive~\cite{Feitelson20142967}. We use the
``cleaned''\cite{pwa-clean} version of the logs as per the archive. We impose
an additional filtering step to the workloads\footnote{This filtering step is
available in the reproducible workflow\cite{repro} as shell script
\lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter|} in order to
clean erroneous data:

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{jsm}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. In our experience, randomly
shuffling months in a trace at random is enough to generate contradictory
conclusions about which algorithm is better. This is a known
problem~\cite{flurries} that motivated decades of research in workload
modeling~\cite{feitbook}. A recent approach~\cite{feitresampling} suggests to
resample data directly from the original traces. In this work, we use a simple
implementation of these ideas.  More precisely, we want to sample the job
submission process, which is achieved in the following manner: (a) For every
week in a resampled trace, we iterate over every existing user in the original
trace; (b) For each of these users, we choose a week of the original trace
uniformly at random; (c) The jobs from this user in this original week are
chosen to be present in the week being resampled.  This simple mechanism
enables us to preserve the dependency within a week.  Although the system state
may not be independent from one week to an other, we rely on existing
work~\cite{jsspp17} and assume for these experiments that the system is in a
stationary state. We do provide guidelines as how to incorporate tracking
into our algorithms in Section~\ref{sec:online} by introducing a decaying
discount factor $\lambda$. Our experiments use stationary resampled data, 
and therefore no decay is necessary: $\lambda$ will be equal to $1$.
